|

### 2023-03-17 manuscript/manuscript_1.0.Rmd
Done with reviewing main text. Need to check figure captions

A somewhat ignored component of communication systems, in the context 
of aggressive interactions, is the cognitive aspects of the receiver module. 
Theoretical models often assume that communication systems relying on badges of 
status to have reaction norms as their 
mechanistic underpinning [@botero_Evolution_2010], with 
individuals using their own quality and the opponent's badge to determine 
whether to aggressively engage in contest. Reaction norms 
allow individuals to respond to the available information without a big
cognitive burden. That is at least in contrast to other information processing 
mechanisms like individual recognition. With individual recognition 
individuals associate cue of their peers to their quality. These association 
must be learned throughout and individual's life, thus it has the usual 
cognitive requirements of an associative learning process. An alternative view 
of badges of status is that individuals learn 
to react to them based on their experiences [@guilford_Receiver_1991], 
which would imply that receivers must learn to associate signals with 
the fighting ability of bearers just as in systems based on individual 
recognition. In cases where badges of status vary quantitatively 
(e.g. in size or intensity), fighting ability may increase monotonically 
with attributes of the badge and would be reinforced by every interaction, 
so in principle learning would be faster than in systems involving individual 
recognition in which the association between signals and their meaning would 
vary depending on the interacting partners. In any case, learning could be 
a central cognitive mechanism in both types of communication systems, 
but the role of learning in these contexts has not been thoroughly 
explored in the empirical nor theoretical literature.

Associative learning is a key cognitive mechanism that allows 
individuals to associate rewards with environmental stimuli 
and thus behave adaptively [@staddon_Adaptive_2016]. Associative 
learning exists in all major vertebrate taxa, and in many invertebrates
as well [@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
@behrens_Associative_2008].
Theory has shown that natural selection favours these associations 
in complex environments where conditions are hard to predict 
[@dridi_Environmental_2016]. Besides its wide taxonomic and
ecological relevance, associative learning is a flexible cognitive 
mechanism whose underpinnings show interspecific variation 
[@enquist_Power_2016; @quinones_Reinforcement_2019]. Associative learning 
is not a single mechanisms, but rather a set of cognitive structures 
and processes that can vary in their scope and complexity. Presumably, 
these structures and processes have been modified 
by natural selection and could provide new explanations for behavioural 
variation. Despite all this, associative learning is 
not often included in the narrative of evolutionary explanations of 
behavioural patterns [@fawcett_Exposing_2013; @kamil_Optimal_1983; 
@mcauliffe_Psychology_2015]. Computational models of evolution 
can play an important role to overcome the lack of integration 
between learning and evolution. Reinforcement learning theory 
encompasses a series of computational methods inspired 
on the psychological and neurological mechanisms of associative 
learning [@sutton_Reinforcement_2018]. These set of algorithms allow the 
implementation of biologically realistic problems, 
capturing the essence of learning processes 
[@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
Furthermore, these algorithms can be embedded in evolutionary simulations to 
generate theoretical predictions of the effect that learning can 
have in behavioural evolution [@leimar_Learning_2019;@leimar_Effects_2022].

In here we present an evolutionary model where individuals use 
associative learning to develop a tendency to behave aggressively 
(or  peacefully), in the context of competition over resources, 
depending on the quantitative morphological trait they 
perceive in their opponent (badge). Over evolutionary time 
individuals evolve the size of their badge, and whether it depends 
on their quality. Under this simple
set up, individuals can use the badge as a signal of quality. 
We use the model to assess under what conditions 
of interaction structure we expect different communication signals to evolve.

@@

We model the evolution of signals of quality in the context of agonistic 
interactions. For simplicity we consider a population of haploid 
individuals with non-overlapping generations. Individuals are born 
every generation with a level of quality ($Q_i$, where $i$ is subscript 
of the population vector of size $N$) given by a number between zero and one, 
which is drawn from a truncated normal distribution 
$N( 0.5,\sigma)$. An individual with quality $0$ has the lowest RHP, 
while an individual with quality $1$ has the highest. As part of 
development processes individuals produce a morphological signal (badge), 
the size of such depends on their level 
of quality according to a reaction norm given by 
$B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$; where $\alpha_i$ and 
$\beta_i$ are individual specific traits that determine the shape of 
the reaction norm (Fig. \ref{fig:model-struc} A). Variation in $\alpha_i$ 
and $\beta_i$ means individuals can have either uninformative (flat) or 
informative (logistic) norms. The size of the signal is constrained 
between zero and one. We assume different values of these traits 
are given by different alleles, and are inherited from mother to offspring. 
Unless, with a small probability ($\mu$), mutation changes the allelic value 
of the offspring by an amount taken from a normal distribution $TN(0,\sigma_\mu)$.
After birth individuals go through a round of 
viability selection. Individual specific survival probability is 
given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$. Where $k_1$ and $k_2$ are parameter 
values determining the shape of the survival function. 
Importantly, if $k_2=0$ survival probability is independent of quality, 
while if $k_2>0$ lower quality individuals pay a higher price for similar-sized 
badge, fulfilling the core assumption of the handicap principle 
[@botero_Evolution_2010; @grafen_Biological_1990; @johnstone_Badges_1993].

Individuals who survive engage in a series of pairwise interactions 
where they compete for resources. 
<!-- Individuals in real populations typically  -->
<!-- do not interact at random nor with all individuals in the population  -->
<!-- [@kurvers_Evolutionary_2014]. In order to allow for non-random interactions,  -->
<!-- we use the population vector to define an interaction neighbourhood  -->
<!-- for the focal individual, the size of the neighbourhood ($g$) -->
<!-- is the number of positions around the focal's from which the partner is drawn.  -->
<!-- So, if $g=N$ interactions are global; as $g$ gets smaller  -->
<!-- interactions are more local.  -->
In an interaction each individual must 
decide whether to escalate a fight or not. 
Following the classic *hawk-dove* game [@maynard-smith_Evolution_1982], 
if the focal individual fights and its partner does not, the focal gets 
as pay-off the resource of value $V$ while its partner gets nothing; 
if both individuals restrain  from fighting they split up the resource 
in half; if both decide to fight they split up the cost of fighting and 
the winner takes over the resource. We further assume that the 
probability of wining a fight for the focal individual depends on the 
difference in quality between her and her interacting partner; specifically 
it is given by $p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})$, where $k_3$ is a 
parameter defining how strong the quality difference determines the 
winning probability; and $i$ and $j$ denote the the position 
in the population vector of the focal and its interacting 
partner respectively.

The decision of whether to escalate a fight against an interacting 
partner can be dependent on the size of the partner's badge, 
the actual dependency is determined by the focal´s experiences 
through a learning process. We implement learning using 
the actor-critic approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
Individuals estimate the reward (pay-off) expected from  interacting 
with partners of different badge sizes (the critic in RL terminology). 
After every interaction they update the estimate of reward proportionally 
to the difference between their current estimate and the observed 
reward (prediction error $\delta$) and to the speed of learning 
($\mathrm{A}$). Furthermore, they express different probabilities of 
retreating (attacking) depending on the badge size of their opponent 
(the actor in RL). They update the probability of 
retreating (attacking)  up or down depending on whether retreating 
(attacking) leads to an increase in the reward estimation. 
So, if a focal individual decides to escalate a fight against 
an individual with small badge, and that leads to an increase in 
the reward estimation, the focal individual will increase the probability 
of escalating fights with individuals of small badges in the future. 
Given that badge size is a real number between 0 and 1, there are infinitely 
many badge sizes. Thus, the reward estimation, as well as the probability 
of retreating (escalating), must be generalized across different values. 
To implement generalization we use the linear function approximation 
method based on radial basis functions [@sutton_Reinforcement_2018]. 
Specifically, we pick $c$ feature centres, which are evenly spaced values 
along the badge size interval ([0,1]). This feature centres are the same for all individuals and stored in vector $\symbf{b}$ . Each one of these features centres 
is associated with a weight for reward estimation and tendency to play 
retreat in a given interaction. The reward estimation and the tendency 
to retreat are calculated (in every interaction) as the sum of the weights 
associated with each feature centre, weighted by the response 
triggered by the feature (Fig. \ref{fig:model-struc} B, dots 
represent the feature weights) The response of each feature centre 
diminishes as a Gaussian function  with the distance between the 
feature centre and the badge size of 
the partner (Fig. \ref{fig:model-struc} B, grey line). Formally the 
reward estimation $\hat{R}$ when the focal individual ($i$) faces 
individual $j$ is given by,
\
\begin{equation}
\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
\end{equation}

@@

After all their interactions, individuals in the population 
reproduce proportionally to their total pay-off $w_i$, which is a sum of the 
baseline pay-off ($w_0$) and all the pay-offs obtained throughout their 
life. Thus, the combination of natural selection and genetic drift 
changes the distribution of values in $\alpha$ and $\beta$ that segregates 
in the population, effectively changing  the badges expressed and the 
communication system.

```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions. In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual. Red shows an informative reaction norm, while the blue shows a uninformative reaction norm. B the receiver code: individuals have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Generalization is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward. In C and D, effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A), while in D they face signallers with informative reaction norms (red line in A). Colour scale in C (applies also for D) indicates the quality of the individual"}

@@

rangx<-seq(0,1,by=0.01)
  dataIndsAct<-sapply(as.list(tempPop[nInteract==behavTime,indId]),
                      function(x){x=
                        logist(totRBF(rangx,
                                      centers,sigSquar,
                                      as.double(
                                        tempPop[(nInteract==behavTime
                                                 &indId==x)
                                                # &(Quality>0.8&Quality<1)
                                                ,.SD,
                                                .SDcol=grep("WeightAct",
                                                            names(tempPop),
                                                            value = TRUE)
                                                ])),alpha = 0,beta = 1)})
  # par(plt=posPlot(numploty = 4,numplotx = 4,idploty = 2,idplotx = 1))
  par(plt=posPlot(numploty = 1,numplotx = 2,idploty = 1,idplotx = 1,upboundx = 91))
matplot(x=rangx,y=dataIndsAct,type='l',xlab="",ylab="prob. retreat",
          xaxt="s",yaxt="s",lty = 1,
          col=paletteMeans(100)[
            findInterval(tempPop[nInteract==1700,Quality],colorbreaksQual)],
          lwd=3,ylim=c(0,1.1))
text(x = 0.08,y=1.05,"B",cex = 1.5)
par(new=FALSE)
color.bar.aeqp(paletteMeans(100),min =min(colorbreaksQual),
               max = max(colorbreaksQual),nticks = 3,
               cex.tit = 0.9,title = "quality",
               numplotx = 20,numploty = 10,idplotx =9,idploty = 2)

@@

We first present the outcome of simulations where we prevent the 
evolution of the sender code (by setting the mutation rate to $0$), 
and assume that all individuals in the population display either 
an uninformative or informative badge size with respect to their quality. 
These simulations show what type of receiver strategy develops through 
a learning process. Figure \ref{fig:model-struc} (C and D) shows 
the receiver strategy developed through learning; panel C is for 
receivers that faced uninformative signals, and D informative ones. 
When learners face uninformative signals, they change their 
probability of retreating depending on their own quality. Individuals 
with higher quality (red tones) after the learning process have a 
high probability of attacking; while individuals of lower quality 
(blue tones) mostly retreat from confrontations. Thus, learning splits
the population of receivers into the two classic pure strategies of 
hawks and doves. Given that we have a assumed a monomorphic population 
with unresponsive reaction norms on the signalling side, the changes 
triggered by learning only affect a small range of badge sizes. 
In contrast, when receivers face informative reaction norms 
on the side of the signaller (panel B in fig. \ref{fig:model-struc}), 
receivers use the badge size of their interacting partners to determine 
whether to retreat or attack. The relationship is given by a 
threshold-like reaction norm, where the change from retreating to 
attacking depends on the quality of the receiver. As expected, 
the higher the quality of the receiver the larger the badge size 
that triggers a retreat.

@@

When we allow the evolution of the sender code ($\alpha$ and $\beta$ 
change subject to natural selection and genetic drift), and the badge 
size work as a handicap (i.e. the cost of the badge is inversely 
proportional to the quality of the individual), the sender code can evolve 
to produce an honest signal (Fig. \ref{fig:handicap}). The evolution of 
the sender code does not happen immediately after the start of the 
evolutionary process. Instead, the evolutionary dynamics of the reaction 
norm parameters (Fig. \ref{fig:handicap}, A) show a set of 
steps. First, the intercept of the reaction norm ($\alpha$) evolves to 
higher values, reducing the average badge size in the populations. 
That makes sense because reaction norms segregating in the population are 
flat at this point, and consequently the receiver code does not respond to 
the badge size, bigger badges are costlier and do not trigger lower attack 
probabilities. During those first generations, the slope of the sender 
reaction norm ($\beta$)  remains around the initial value of zero. 
At around generation 4000, the slope evolves toward positive values. 
Positive values in the slope imply that larger badges correlate 
with higher quality (Fig. \ref{fig:handicap} B). Receivers learn to react to such 
correlation, increasing the probability of retreat when facing 
individual of larger badges (Fig. \ref{fig:handicap} C). Hence, natural selection 
favours larger values of $\beta$, eventually leading to an evolutionary 
equilibrium where badge size is an honest signal of quality mediated
by the learned responses of receivers (Fig. \ref{fig:handicap} B.4 and C.4). 
The evolutionary trajectory portrayed in figure 
\ref{fig:handicap} A is not the only possible trajectory. If the slope of the 
sender reaction norms ($\beta$) evolves toward negative values before badges 
become handicaps, receivers never learn to react to the size of badges. 
Thus, badges are only costly and do not provide information about the 
quality of an individual (Fig. \ref{fig:noBadge}) . Eventually, the badge 
disappears from the population. In contrast, when the badge does not 
work as a handicap but instead is cost-free, the evolutionary process 
never leads to the establishment of an honest signal. 
Instead, subject to genetic drift the 
population either evolves towards the disappearance of the badge, or to the 
maximum badge size. In either of this cases reaction norms are flat, so the 
badge does not provide any quality information.

The amount of information that agents are able to collect through learning 
throughout their life time can strongly change the outcome of 
the evolutionary dynamics. In the simulations presented in figure 
\ref{fig:handicap} and \ref{fig:noBadge}, 
individuals learned with high learning rate ($\alpha = 0.4$) and interacted 
repeatedly along their lifetime (on average 2000 interactions). When we reduce 
the number of interactions that individuals have over their lifetime to 300 
interactions on average, we see an drastic increase in the morphological and genetic
variation present in the population. Fig. \ref{fig:branching} shows 
changes in the distribution of the sender reaction norm parameters
along evolutionary time, assuming the badge is cost-free. 
In panels A1 and 2, darker areas show trait values with high frequency in the population. 
Populations start monomorphic with a value of zero on both traits, 
and mutations quickly build up a normal distribution around the starting 
value. Within the first 2000 generations the unimodal distribution in 
the intercept ($\alpha$) splits into a bimodal one. 
Later in evolutionary time one of the peaks splits further, 
so at the end of the evolutionary simulation the distribution of the intercept 
($\alpha$) in the population shows three peaks. In the case of $\beta$, the 
peaks in the distribution are not so clear-cut, but it's clear that there is an 
increase in the variance of the distribution. These changes in the 
distribution of the sender reaction norm parameters 
imply that individuals can generally be classified into three 
distinct types (\ref{fig:branching}). Two types express a flat reaction 
norm with extreme values for the badge size, meaning that their badge size 
is not informative of their quality, while the third type shows 
intermediate badge sizes which are determined 
by the quality of the individual (Fig. \ref{fig:branching} B.3-4). 
In this particular simulation, the values of the slope in the informative group
are negative, which implies that badge size correlates negatively with quality. 
Furthermore, the receiver reaction norms developed through learning, particularly those 
of individuals of intermediate quality, respond to the signal of their 
sender type by increasing the probability of retreating from a fight with 
individuals with smaller badges (Fig. \ref{fig:branching}).

>>>>

A somewhat ignored component of communication systems in the context 
of aggressive interactions are cognitive aspects of the receiver module. 
Theoretical models often assume that communication systems relying on badges of 
status have reaction norms as their mechanistic underpinning 
[@botero_Evolution_2010], with 
individuals using their own quality and the opponent's badge to determine 
whether to aggressively engage in contests. Reaction norms 
allow individuals to respond to the available information without a large
cognitive burden. The computations involved in the reaction norm response do not
required a lot of memory or advanced algorithmic processes. 
This is in contrast to other information processing 
mechanisms like individual recognition. With individual recognition, 
individuals associate cues of their peers to their quality. Because these 
associations must be learnt throughout the life of individual, 
they have the usual cognitive requirements of associative learning processes. 
An alternative view of badges of status is that individuals learn 
to react to them based on their experiences [@guilford_Receiver_1991], 
which would imply that receivers must learn to associate signals with 
the fighting ability of bearers just as in systems based on individual 
recognition. In cases where badges of status vary quantitatively 
(e.g. in size or intensity), fighting ability may increase monotonically 
with attributes of the badge and would be reinforced by every interaction.
Therefore, in principle, learning in such scenarios would be faster 
than in systems involving individual 
recognition in which the association between signals and their meaning 
varies depending on the interacting partners. In any case, learning may be 
a central cognitive mechanism in both types of communication systems, 
but the role of learning in these contexts has not been thoroughly 
explored in the empirical nor theoretical literature.

Associative learning is a taxonomically widespread cognitive 
mechanism that allows individuals to associate rewards with environmental stimuli 
and thus behave adaptively 
[@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
@behrens_Associative_2008].
Theory has shown that natural selection favours such associations 
in complex environments where conditions are difficult to predict 
[@dridi_Environmental_2016]. Associative learning is a flexible cognitive 
mechanism whose underpinnings show interspecific variation 
[@enquist_Power_2016; @quinones_Reinforcement_2019]. Associative learning 
is not a single mechanisms, but rather a set of cognitive structures 
and processes that can vary in their scope and complexity. Presumably, 
these structures and processes have been modified 
by natural selection and could provide new explanations for behavioural 
variation. Despite all this, associative learning is 
not often included in the narrative of evolutionary explanations of 
behavioural patterns [@fawcett_Exposing_2013; @kamil_Optimal_1983; 
@mcauliffe_Psychology_2015]. Computational models of evolution 
can play an important role to overcome the lack of integration 
between learning and evolution. Reinforcement learning theory 
encompasses a series of computational methods inspired 
on the psychological and neurological mechanisms of associative 
learning [@sutton_Reinforcement_2018]. This set of algorithms allows the 
implementation of biologically realistic problems, 
capturing the essence of learning processes 
[@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
Furthermore, these algorithms can be embedded in evolutionary simulations to 
generate theoretical predictions of the effect that learning can 
have in behavioural evolution [@leimar_Learning_2019;@leimar_Effects_2022].

Here we present an evolutionary model where individuals use 
associative learning to develop a tendency to behave aggressively 
(or  peacefully) in the context of competition over resource, 
depending on a quantitative morphological trait 
(i.e.: a badge indicating quality) they 
perceive in their opponents. Over evolutionary time the size of 
the badge evolves, and its dependency on the individual's quality. 
Under this simple set up, individuals can use the badge as a signal of quality. 
We use the model to assess under what conditions 
we expect communication signals to evolve as handicaps or conventions.

@@

We model the evolution of signals indicating individual quality 
in the context of agonistic interactions. 
For simplicity we consider a population of haploid 
individuals with non-overlapping generations. Individuals are born 
every generation with a level of quality ($Q_i$, where $i$ is subscript 
of the population vector of size $N$) given by a number between zero and one, 
which is drawn from a truncated normal distribution 
$N( 0.5,\sigma)$. An individual with quality $0$ has the lowest RHP, 
while an individual with quality $1$ has the highest. As they 
develop, individuals produce a phenotypic signal (badge), 
the size of which depends on their quality according to a reaction norm given by 
$B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$; where $\alpha_i$ and 
$\beta_i$ are individual specific traits that determine the shape of 
the reaction norm (Fig. \ref{fig:model-struc} A). Variation in $\alpha_i$ 
and $\beta_i$ means individuals can have either uninformative (flat) or 
informative (logistic) norms. Informative reaction norms represent a 
developmental program where the phenotype of the individual is determined by
the environmental conditions under which it grows, thus causing a 
correlation between quality and signal. The size of the signal is constrained 
to take values between zero and one. We assume different values of these traits 
are given by different alleles and are inherited from mother to offspring 
unless, with a small probability ($\mu$), mutation changes the allelic value 
of the offspring by an amount taken from a normal distribution $TN(0,\sigma_\mu)$.

After birth, individuals go through a round of 
viability selection. The survival probability of an individual is 
given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$, where $k_1$ and $k_2$ are parameter 
values determining the shape of the survival function. 
Importantly, if $k_2=0$ survival probability is independent of quality, 
while if $k_2>0$ lower quality individuals pay a higher price for similar-sized 
badge, fulfilling the core assumption of the handicap principle 
[@botero_Evolution_2010; @grafen_Biological_1990; @johnstone_Badges_1993].

Individuals who survive engage in a series of pairwise interactions 
where they compete for resources. 
<!-- Individuals in real populations typically  -->
<!-- do not interact at random nor with all individuals in the population  -->
<!-- [@kurvers_Evolutionary_2014]. In order to allow for non-random interactions,  -->
<!-- we use the population vector to define an interaction neighbourhood  -->
<!-- for the focal individual, the size of the neighbourhood ($g$) -->
<!-- is the number of positions around the focal's from which the partner is drawn.  -->
<!-- So, if $g=N$ interactions are global; as $g$ gets smaller  -->
<!-- interactions are more local.  -->
In each interaction individuals must 
decide whether to escalate a fight or not. 
Following the classic *hawk-dove* game [@maynard-smith_Evolution_1982], 
if the focal individual fights and its partner does not, the focal gets 
as pay-off the resource of value $V$ while its partner gets nothing; 
if both individuals restrain  from fighting, then they split up the resource 
in half; if both decide to fight they split up the cost of fighting and 
the winner takes over the resource. We further assume that the 
probability of wining a fight for the focal individual depends on the 
difference in quality between it and its interacting partner; specifically 
it is given by $p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})$, where $k_3$ is a 
parameter defining how strong the quality difference determines the 
winning probability, and $i$ and $j$ denote the the position 
in the population vector of the focal and its interacting 
partner, respectively.

The decision of whether to escalate a fight against an interacting 
partner can depend on the size of the partner's badge, 
the dependency is determined by the focal's experiences 
through a learning process. We implement learning using 
the actor-critic approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
Individuals estimate the reward (pay-off) expected from  interacting 
with partners of different badge sizes (the critic in RL terminology). 
After each interaction they update the estimate of reward proportionally 
to the difference between their current estimate and the observed 
reward (prediction error $\delta$) and to the speed of learning 
($\mathrm{A}$). Furthermore, individuals express different probabilities of 
retreating/attacking depending on the badge size of their opponent 
(the actor in RL). They update the probability of 
retreating/attacking  up or down depending on whether retreating/attacking
leads to an increase in the reward estimation. 
So, if a focal individual decides to escalate a fight against 
an individual with small badge and this leads to an increase in 
the reward estimation, then the focal individual will increase the probability 
of escalating fights with individuals of small badges in the future. 
Given that badge size is a real number between 0 and 1, there are infinitely 
many badge sizes. Thus, the reward estimation, as well as the probability 
of retreating/attacking, must be generalized across different values. 
To implement generalization we use the linear function approximation 
method based on radial basis functions [@sutton_Reinforcement_2018]. 
Specifically, we pick $c$ feature centres, which are evenly spaced values 
along the badge size interval ([0,1]) where the updates are focused. 
For the sake of simplicity we keep the location of these feature centres 
constant and  the same for all individuals, the are stored in the 
vector $\symbf{b}$ . Each of these feature centres 
is associated with a weight for reward estimation and tendency to play 
retreat in a given interaction. The reward estimation and the tendency 
to retreat are calculated (in every interaction) as the sum of the weights 
associated with each feature centre, weighted by the response 
triggered by the feature (Fig. \ref{fig:model-struc} B, dots 
represent the feature weights). The response of each feature centre 
diminishes as a Gaussian function with the distance between the 
feature centre and the badge size of 
the partner (Fig. \ref{fig:model-struc} B, grey line). Formally, the 
reward estimation $\hat{R}$ when the focal individual ($i$) faces 
individual $j$ is given by,
\
\begin{equation}
\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
\end{equation}

@@

Individuals interact as the focal individual $n$ times, and the interaction 
partner is chosen at random. Thus, the expected number of interactions for 
each individual is $n+\frac{n(N+1)}{N}$. 
After the interaction round , individuals in the population 
reproduce with a probability proportional to their total pay-off $w_i$, which is a sum of the 
baseline pay-off ($w_0$) and all the pay-offs obtained throughout their 
life. Thus, the combination of natural selection and genetic drift 
changes the distribution of values in $\alpha$ and $\beta$ that segregates 
in the population, effectively changing  the badges expressed and the 
communication system.

```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions. In A, the reaction norm determines the badge size. Red shows an informative reaction norm, while the blue shows a uninformative reaction norm. B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Feature weaights are values associated with a particular badge size and are updated as individuals interact with each other. The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward. In C and D, effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A), while in D they face signallers with informative reaction norms (red line in A). Colour scale in C and D indicates the quality of the individual"}

@@

rangx<-seq(0,1,by=0.01)
  dataIndsAct<-sapply(as.list(tempPop[nInteract==behavTime,indId]),
                      function(x){x=
                        logist(totRBF(rangx,
                                      centers,sigSquar,
                                      as.double(
                                        tempPop[(nInteract==behavTime
                                                 &indId==x)
                                                # &(Quality>0.8&Quality<1)
                                                ,.SD,
                                                .SDcol=grep("WeightAct",
                                                            names(tempPop),
                                                            value = TRUE)
                                                ])),alpha = 0,beta = 1)})
  # par(plt=posPlot(numploty = 4,numplotx = 4,idploty = 2,idplotx = 1))
  par(plt=posPlot(numploty = 1,numplotx = 2,idploty = 1,idplotx = 1,upboundx = 91))
matplot(x=rangx,y=dataIndsAct,type='l',xlab="",ylab="prob. retreat",
          xaxt="s",yaxt="s",lty = 1,
          col=paletteMeans(100)[
            findInterval(tempPop[nInteract==1700,Quality],colorbreaksQual)],
          lwd=3,ylim=c(0,1.1))
text(x = 0.08,y=1.05,"C",cex = 1.5)
par(new=FALSE)
color.bar.aeqp(paletteMeans(100),min =min(colorbreaksQual),
               max = max(colorbreaksQual),nticks = 3,
               cex.tit = 0.9,title = "quality",
               numplotx = 20,numploty = 10,idplotx =9,idploty = 2)

@@

We first present the outcome of simulations where we prevent the 
evolution of the badge size (by setting the mutation rate to $0$), 
and assume that all individuals in the population display either 
an uninformative or informative badge size with respect to their quality. 
These simulations show what type of receiver strategy develops through 
a learning process. Figure \ref{fig:model-struc} (C and D) shows 
the receiver strategy developed through learning; panel C is for 
receivers that faced uninformative signals, and D informative ones. 
The simulations reveal that when learners face uninformative signals, 
they modify their probability of retreating depending on their own quality. 
Individuals with higher quality (red tones) after the learning process have a 
high probability of attacking; while individuals of lower quality 
(blue tones) mostly retreat from confrontations; in both cases irrespective of 
their opponent. Thus, learning splits the population of receivers into 
the two classic pure strategies of hawks and doves. 
Given that we have a assumed a monomorphic population 
with unresponsive reaction norms on the signalling side, the changes 
triggered by learning only affect a small range of badge sizes. This can be seen 
in panel C, all learning has occurred around a badge size of $0.5$, because all
badges in the population are this size. 
In contrast, when receivers face informative reaction norms 
on the side of the signaller (panel B in fig. \ref{fig:model-struc}), 
receivers use the badge size of their interacting partners to determine 
whether to retreat or attack. The relationship is given by a 
threshold-like reaction norm, where the decision to retreat or  
attack depends on the quality of the receiver. As expected, 
the higher the quality of the receiver the larger the badge size 
that triggers a retreat.

@@

When we allow the badge size to evolve ($\alpha$ and $\beta$ 
change subject to natural selection and genetic drift), and the signal
works as a handicap (i.e. the cost of the badge is inversely 
proportional to the quality of the individual), the sender code can evolve 
to produce an honest signal (Fig. \ref{fig:handicap}). The evolution of 
the sender code does not happen immediately after the start of the 
evolutionary process. Instead, the evolutionary dynamics of the reaction 
norm parameters (Fig. \ref{fig:handicap}, A) appear to involve a set of 
steps. First, the intercept of the reaction norm ($\alpha$) 
describing the relationship between individual quality and badge size evolves to 
higher values, reducing the average badge size in the populations. 
This is because reaction norms segregating in the population are 
flat at the beginning of the simulation, consequently, 
the receiver code does not respond to the badge size, 
and larger badges are costlier and do not trigger lower attack 
probabilities. During the first generations, the slope of the sender 
reaction norm ($\beta$)  remains close to the initial value of zero. 
At around generation 4000, the slope evolves toward positive values. 
Positive values in the slope imply that larger badges correlate 
with higher quality (Fig. \ref{fig:handicap} B). Receivers then learn to react to such 
correlation, increasing the probability of retreat when facing 
individuals with larger badges (Fig. \ref{fig:handicap} C). Hence, natural selection 
favours larger values of $\beta$, eventually leading to an evolutionary 
equilibrium in which badge size is an honest signal of quality mediated
by the learned responses of receivers (Fig. \ref{fig:handicap} B.4 and C.4).

The evolutionary trajectory portrayed in figure 
\ref{fig:handicap} A is not the only possible outcome. If the slope of the 
sender reaction norms ($\beta$) evolves, subject to genetic drift,
toward negative values before badges become handicaps, 
receivers never learn to react to the size of badges. 
Thus, badges are only costly and do not provide information about the 
quality of individuals (Fig. \ref{fig:noBadge}) . Eventually, the badge 
disappears from the population. In contrast, when the badge does not 
work as a handicap but instead is cost-free, the evolutionary process 
never leads to the establishment of an honest signal. 
Instead, subject to genetic drift, the 
population either evolves towards the disappearance of the badge, or to the 
maximum badge size. In either of this cases reaction norms are flat, so the 
badge does not provide any information about quality.

The amount of information that agents are able to collect through learning 
throughout their life time can strongly change the outcome of 
evolutionary dynamics. In the simulations presented in figure 
\ref{fig:handicap} and \ref{fig:noBadge}, 
individuals learned with high speed of learning ($\mathrm{A} = 0.4$) and interacted 
repeatedly along their lifetime (on average 2000 interactions). When we reduced 
the number of interactions that individuals have over their lifetime to 300 
on average, we saw a drastic increase in the phenotypic and genetic
variation present in the population. Fig. \ref{fig:branching} shows 
changes in the distribution of the sender reaction norm parameters
along evolutionary time, assuming the badge is cost-free. 
In panels A1 and 2, darker areas show trait values with high frequency in the population. 
Populations start monomorphic with a value of zero on both the intercept and
the slope, and mutations quickly build up a normal distribution around the starting 
value. Within the first 2000 generations, the unimodal distribution in 
the intercept ($\alpha$) splits into a bimodal one. 
Later in evolutionary time one of the peaks splits further, 
so at the end of the evolutionary simulation the distribution of the intercept 
($\alpha$) in the population shows three peaks. In the case of $\beta$, the 
peaks in the distribution are not so clear-cut, but it is clear that there is an 
increase in the variance of the distribution. These changes in the 
distribution of the parameters of the sender reaction norm 
imply that individuals can generally be classified into three 
distinct types (\ref{fig:branching}). Two types express a flat reaction 
norm with extreme values for the badge size, meaning that their badge size 
is not informative of their quality, whereas the third type shows 
intermediate badge sizes determined by the quality of the individual 
(Fig. \ref{fig:branching} B.3-4). 
In this particular simulation, the values of the slope in the informative group
are negative, which implies that badge size correlates negatively with quality. 
Furthermore, the receiver reaction norms developed through learning, particularly those 
of individuals of intermediate quality, respond to the signal of their 
sender type by increasing the probability of retreating from a fight with 
individuals with smaller badges (Fig. \ref{fig:branching}).
